Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Saturn-Sized Planet in IC~2602

### 2020-07-09 paper/ms.tex
update manuscript
(non-converged chains, but looking decent)

Over the first 100 million years of their lives, exoplanet systems are
expected to undergo major physical and dynamical changes.  For a
typical Sun-like star, the protoplanetary disk disperses within
roughly 1--10 million years
\citep{mamajek_initial_2009,dullemond_inner_2010,williams_protoplanetary_2011}.
Gas giants presumably finish accreting before the end of disk
dispersal \citep{pollack_formation_1996}.  While rocky planets may
fully form within only a few million years
\citep{dauphas_hf-w-th_2011}, they can also undergo significant
changes over the next $\sim$10-100 million years through giant impacts
\citep[{\it
e.g.},][]{kleine_hf-w_2009,konig_earths_2011,raymond_terrestrial_2014}.
The Moon, for instance, may have formed from debris ejected during a
collision between the proto-Earth and a planetesimal during Earth's
first 100 million years \citep{cameron_origin_1976,canup_origin_2001}.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f2a.pdf}
		}

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{
    {\bf Photometry of \tn.} {\it Top:}
    Phase-folded TESS transit. Gray points are 2-minute PDCSAP flux
    measurements with the local spot-induced variation removed.  Black
    points are binned to 6-minute intervals. The black error-bar shows
    the median uncertainty for the black points.  The gray line is the
    best-fit model to the combined TESS and ground-based data
    (Table~\ref{tab:posterior}).  {\it Bottom:} Multicolor photometry
    from the El Sauce 0.36m; black points are binned to 10-minute
    intervals.  Models show $2\sigma$ lower limits on the transit
    depths in the Cousins-R and Johnson-B bandpasses used to rule out
    specific false positive scenarios (see
    Section~\ref{subsec:colorphot}).
    \label{fig:jointphot}
	}
\end{figure*}

@@

Figure~\ref{fig:rawzoom} shows the data.  The dominant modulation
induced by starspots coming into and out of view has a peak-to-peak
amplitude of about 2.3\%, and a period of about 3 days.  The dips
suggestive of a grazing transiting planet recur roughly every 8 days,
and have a depth of about 0.4\%.  A few flares are visible.  A
phase-folded view of the TESS transits with starspot variability
removed is shown in Figure~\ref{fig:jointphot}. Our fitting procedure
for the models shown in these plots is described in
Section~\ref{subsec:planet}.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.85\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Astrophysical false positive scenarios.}
    {\it Top}: for bound companions (EB and HEB scenarios),
		{\it Bottom}: for unassociated companions along the same line of
    sight (BEB scenarios).
		Each constraint is described in Section~\ref{subsec:fp_constraints}.
    Gray regions are ruled out by at least one constraint.
		\label{fig:fpscenario}
	}
\end{figure*}

@@

We fitted for the observed depths in different bandpasses using a
similar machinery as described in Section~\ref{subsec:planet}.  We
fitted each ground-based transit individually for the planet-to-star
size ratio, the impact parameter, and a local quadratic trend (the
ephemeris was assumed from an initial fit of only the TESS data).  The
corresponding $2$-$\sigma$ limits on the transit depths in Cousins-R
and Johnson-B band light curves were 2.82 and 1.77$\,$ppt,
respectively, and are shown in Figure~\ref{fig:jointphot}.
Particularly in our Johnson-B light curve, the transit depth is
correlated with the mean and linear slope of the light curve: a
smaller depth is allowed if the data are fitted with a larger linear
slope and a larger mean.  Our quoted limits marginalize over these
correlations, and the depth measurement itself is nearly Gaussian.

@@

% results from 20200624_simple_fpp_weak_sec 
We ran \texttt{VESPA}\footnote{We used \texttt{VESPA-0.6} and
\texttt{isochrones-1.2.2}.}, and directly incorporated our constraints
of the SOAR $I$-band contrast curve and a non-detection of secondary
eclipses with a depth set at roughly twice the limits from the SPOC
vetting report (0.1\%).  We verified that changing the secondary
eclipse depth limit did not significantly affect the results.  We set
the maximum aperture radius at $2''$, based on our ground-based
photometry.  Incorporating the constraints from
Figure~\ref{fig:fpscenario}, our nominal false positive probability
analysis excluded EB and HEB scenarios.  This yielded an FPP of 0.21\%
for \pn, sufficient for formal validation as a planet
\citep{morton_efficient_2012}.  We did not incorporate our constraint
that \tn\ is not double-lined, which rules out an additional portion
of BEB parameter space.  Had we not acquired multicolor ground-based
photometry, and been unable to exclude HEB scenarios, the FPP would
have risen to 8\%.  Since the transits are achromatic
(Figure~\ref{fig:jointphot}), particularly in Johnson-B band, we can
rule out HEB scenarios.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.75\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
  {\bf Hertzsprung-Russell diagram of \tn\ and members of \cn.}
  Members (black circles) were identified by
  \citet{cantatgaudin_gaia_2018}.  Gray circles are non-member stars
  within 5 standard deviations of the mean \cn\ right ascension,
  declination, and parallax.  $G$ denotes Gaia broadband magnitudes,
  $Bp$ Gaia blue, and $Rp$ Gaia red.  MIST isochrones (colored lines)
  fit the upper main sequence well, but diverge from the data for
  $M_\star \lesssim 0.7 M_\odot$. This is a known problem with PMS
  models (see Section~\ref{subsec:hr}).
  \label{fig:hr}
	}
\end{figure*}

@@

\input{starparams.tex}

@@

\input{model_posterior_table.tex}

@@

Our joint model for the TESS and ground-based data was an
\citet{exoplanet:agol19} transit with physical and orbital parameters
shared across all transit windows, plus a local quadratic trend
allowed within each window.  The parameters and priors
are listed in Table~\ref{tab:posterior}.  In brief, we fitted for the
shared stellar parameters ($\log g, R_\star, u_0, u_1$), and the
shared planetary parameters ($t_0, P, b, \log(R_{\rm p}/R_\star)$).
There were also three free trend parameters for each of the nine
transits to account for the rotational variability.  This yielded a
total of 35 free parameters, of which 8 have physical relevance, and
27 are nuisance parameters.

@@

% \clearpage
% \clearpage

\input{TOI837_phot_table.tex}

\input{TOI837_rv_table.tex}

>>>>

Over the first 100 million years of their lives, exoplanet systems are
expected to undergo major physical and dynamical changes.  For a
typical Sun-like star, the protoplanetary disk disperses within
roughly 1--10 million years
\citep{mamajek_initial_2009,dullemond_inner_2010,williams_protoplanetary_2011}.
Gas giants presumably finish accreting before the end of disk
dispersal \citep{pollack_formation_1996}.  While rocky planets may
fully form within only a few million years
\citep{dauphas_hf-w-th_2011}, they can also undergo significant
growth over the next $\sim$10-100 million years through giant impacts
\citep[{\it
e.g.},][]{kleine_hf-w_2009,konig_earths_2011,morbidelli_building_2012,raymond_terrestrial_2014}.
The Moon, for instance, may have formed from debris ejected during a
collision between the proto-Earth and a planetesimal during Earth's
first 100 million years \citep{cameron_origin_1976,canup_origin_2001}.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.47\textwidth]{f2a.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Phase-folded TESS transit.} Gray points are 2-minute PDCSAP flux
    measurements with the local spot-induced variation removed.  Black
    points are binned to 6-minute intervals. The black error-bar shows
    the median uncertainty for the black points.  The gray line is the
    best-fit model to the combined TESS and ground-based data
    (Table~\ref{tab:posterior}). 
		\label{fig:phasefold}
	}
\end{figure}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.47\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Ground-based follow-up photometry.} The data were acquired
    using the El Sauce 0.36m and ASTEP400 (0.40m). Black points are
    binned to 10-minute intervals.  The gray line is the best-fit
    model to the combined TESS and ground-based data. Models show
    $2\sigma$ lower limits on the transit depths in the Cousins-R and
    Johnson-B bandpasses used to rule out specific false positive
    scenarios (see Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

Figure~\ref{fig:rawzoom} shows the data.  The dominant modulation
induced by starspots coming into and out of view has a peak-to-peak
amplitude of about 2.3\%, and a period of about 3 days.  The dips
suggestive of a grazing transiting planet recur roughly every 8 days,
and have a depth of about 0.4\%.  A few flares are visible.  A
phase-folded view of the TESS transits with starspot variability
removed is shown in Figure~\ref{fig:phasefold}. Our fitting procedure
for the models shown in these plots is described in
Section~\ref{subsec:planet}.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Astrophysical false positive scenarios.}
    {\it Top}: for bound companions (EB and HEB scenarios),
		{\it Bottom}: for unassociated companions along the same line of
    sight (BEB scenarios).
		Each constraint is described in Section~\ref{subsec:fp_constraints}.
    Gray regions are ruled out by at least one constraint.
		\label{fig:fpscenario}
	}
\end{figure*}

@@

We fitted for the observed depths in different bandpasses using a
similar machinery as described in Section~\ref{subsec:planet}.  We
fitted each ground-based transit individually for the planet-to-star
size ratio, the impact parameter, and a local quadratic trend (the
ephemeris was assumed from an initial fit of only the TESS data).  The
corresponding $2$-$\sigma$ limits on the transit depths in Cousins-R
and Johnson-B band light curves were 2.82 and 1.77$\,$ppt,
respectively, and are shown in Figure~\ref{fig:groundphot}.
Particularly in our Johnson-B light curve, the transit depth is
correlated with the mean and linear slope of the light curve: a
smaller depth is allowed if the data are fitted with a larger linear
slope and a larger mean.  Our quoted limits marginalize over these
correlations, and the depth measurement itself is nearly Gaussian.

@@

% results from 20200624_simple_fpp_weak_sec 
We ran \texttt{VESPA}\footnote{We used \texttt{VESPA-0.6} and
\texttt{isochrones-1.2.2}.}, and directly incorporated our constraints
of the SOAR $I$-band contrast curve and a non-detection of secondary
eclipses with a depth set at roughly twice the limits from the SPOC
vetting report (0.1\%).  We verified that changing the secondary
eclipse depth limit did not significantly affect the results.  We set
the maximum aperture radius at $2''$, based on our ground-based
photometry.  Incorporating the constraints from
Figure~\ref{fig:fpscenario}, our nominal false positive probability
analysis excluded EB and HEB scenarios.  This yielded an FPP of 0.21\%
for \pn, sufficient for formal validation as a planet
\citep{morton_efficient_2012}.  We did not incorporate our constraint
that \tn\ is not double-lined, which rules out an additional portion
of BEB parameter space.  Had we not acquired multicolor ground-based
photometry, and been unable to exclude HEB scenarios, the FPP would
have risen to 8\%.  Since the transits are achromatic
(Figure~\ref{fig:groundphot}), particularly in Johnson-B band, we can
rule out HEB scenarios.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ 
  {\bf Hertzsprung-Russell diagram of \tn\ and members of \cn.}
  Members (black circles) were identified by
  \citet{cantatgaudin_gaia_2018}.  Gray circles are non-member stars
  within 5 standard deviations of the mean \cn\ right ascension,
  declination, and parallax.  $G$ denotes Gaia broadband magnitudes,
  $Bp$ Gaia blue, and $Rp$ Gaia red.  MIST isochrones (colored lines)
  fit the upper main sequence well, but diverge from the data for
  $M_\star \lesssim 0.7 M_\odot$. This is a known problem with PMS
  models (see Section~\ref{subsec:hr}).
  \label{fig:hr}
	}
\end{figure}

@@

%

@@

%

@@

Our joint model for the TESS and ground-based data was an
\citet{exoplanet:agol19} transit with physical and orbital parameters
shared across all transit windows, plus a local quadratic trend
allowed within each window.  Select parameters and priors
are listed in Table~\ref{tab:posterior}.  In brief, we fitted for the
shared stellar parameters ($\log g, R_\star, u_0, u_1$), and the
shared planetary parameters ($t_0, P, b, \log(R_{\rm p}/R_\star)$).
There were also three free trend parameters for each of the twelve
transits to account for the rotational variability.  This yielded a
total of 44 free parameters, of which 8 have physical relevance, and
36 are nuisance parameters.  To keep the table at a manageable length,
we only report the physically relevant parameters and derived
quantities.

@@

% \clearpage
%\clearpage
\input{starparams.tex}
%\clearpage
\input{model_posterior_table.tex}
\clearpage
\input{TOI837_phot_table.tex}
\input{TOI837_rv_table.tex}
\clearpage
